\chapter*{Abstract}

%This is a summary of the thesis including the conclusion of what was discovered.
Distributed lotteries on the internet is an interesting application with uses in gambling and other areas. Blockchains that function as smart contract platforms can be used to both transfer value and enforce protocols for multiparty computing without relying on a trusted intermediary. This has made it possible to design lotteries with verifiable randomness and with a guarantee of successful completion. 

One promising design of a distributed lottery on a blockchain is based on a tournament of digital coin tosses. This thesis explores the feasibility of such a lottery through making an implementation and doing measurements. The implementation is made for the Ethereum blockchain, which is currently the leading platform for applications using smart contracts. The lottery is assessed by measuring transaction costs and transaction demand, as well as by discussing the security of the lottery in the context of known security issues for blockchain applications and web applications.

We successfully implement a lottery prototype which is likely to work for up to about 100000 participants. Several directions of further research to increase the scalability are identified and discussed. We find that the most concerning security issue is transaction censorship by a powerful collusion of opportunistic miners, which might be an issue for lotteries with a very large prize.

\hypersetup{pageanchor=false}