\section{Ticket price}
\label{sec:ticket-price}

We operate with a gas price of 1.5 gwei, which at the time of writing is considered quite low, but still enough to make the transaction be included in about 30 minutes. Gas prices in Ethereum are denominated in \emph{wei}, which is the most granular unit of ether. $1 wei=1 \mathrm{e}{-18} ether$. A gwei is simply a gigawei ($\mathrm{e}{9} wei$). When quoting prices in USD we operate with an ether price of 176 USD. Most common casino games have a house edge of about 1-10\% \cite{walsh_houses_nodate}, while large lotteries often have a higher house edge of 40-50\% \cite{shackleford_house_nodate}. We therefore consider the organizer taking a 10\% fee of the deposits reasonable for both the players and the organizer. The organizer risks not enough players joining the lottery and thus losing all the ether needed to set up the lottery. 

In our design, a 10\% organizer fee and a 1.5 gwei gas price puts the minimum viable price of each ticket at about 0.025 ether, which at the time of writing is about 5 USD. That's a price we consider acceptable, even though this can change by at least an order of magnitude within months, as the ether price fluctuates a lot. 

For the second design on average, a player spends 82616 gas to join, and for each level, assuming they commit and reveal, 120157 gas. A tournament has $N=2^L$ players and $L$ levels. We index levels of the tournament starting from the root where $l=0$, and up to the first level where $l=L-1$. For a player at each level, the probability of winning is $\frac{1}{2^{l+1}}$. If we assume the prize is 100\% of the deposits, the prize will be $price \cdot N=price \cdot 2^L$. The first lottery level is $L-1$. For it to be worth playing the first match, the price must be so high that $prize \cdot \frac{1}{2^L} > cost_{tx}(L-1)$ or $2^L \cdot price \cdot \frac{1}{2^L} > cost_{tx}(L-1)$. The inequality simply solves as $price > cost_{tx}(L-1)$, which means that if the gas price is 1.5 gwei and $L=10$, the ticket cost must be at least $120157 \cdot 10 \cdot 1.5 \mathrm{e}{-9}=0.0018 ether$ or 0.32 USD. 

This means that the cost of initializing the lottery, which is paid by the lottery organizer, is by far the most significant cost. If we use off-chain negotiation in the matches, the transaction cost will approximately be cut in half, as we only need one transaction from the winner for each match. When calculating a minimum viable price for participating in the lottery, we should therefore primarily consider the price to set up the lottery for the organizer and how much of the ticket income the organizer takes as a fee.
